---
title: "Homework 1"
author: Suning Zhao
Date: September 2022
output: github_document
---
# Loading packages
```{r}
# Set echo = TRUE as default
knitr::opts_chunk$set(echo = TRUE)
# Open the library of tidyverse
library(tidyverse)
```

# Problem 1
##### Download the package containing the penguins dataset and load:

```{r}
# Load the datafile
data("penguins", package = "palmerpenguins")
# Demonstrate the dataset
head(penguins)
# Describe and Summary the dataset
names(penguins)
summary(penguins)
# Compute row, column and mean flipper length
row_penguins = nrow(penguins)
col_penguins = ncol(penguins)
mean_flipper_length = mean(pull(penguins, flipper_length_mm), na.rm = TRUE)
```

##### Write a short description of penguins dataset

*   In _penguins_ dataset, there are 8 variables, including species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex and year. Sex is a factor vectors including 2 different values. Species and island are two factor vectors including 3 different values. Bill_length_mm(mean 43.92, median 44.45, NA 2), bill_depth_mm(mean 17.15, median 17.30, NA 2) are two numeric variables. Flipper_length_mm(mean 200.9, median 197.0, NA 2), body_mass_g(mean 4202, median 4050, NA 2) and year(min 2007, max 2009) are three int variables.


*   The size of _penguins_ is `r row_penguins` rows and `r col_penguins` lines.

*   The mean _flipper_ _length_ is `r mean_flipper_length`mm.



##### Make a scatterplot

```{r}
ggplot(penguins,aes(y = flipper_length_mm, x = bill_length_mm, color = species)) + geom_point()

ggsave("scatter_plot.pdf", height = 4, width = 6)

```

# Problem 2

##### Create a data frame

```{r}
problem2_df = tibble(
# A random sample of size 10 from a standard Normal distribution
   vec_sample = rnorm (10),
# A logical vector indicating whether elements of the sample are greater than 0
   vec_logical = vec_sample > 0,
# A character vector of length 10
   vec_char = c("A", "B", "C", "D", "E","F", "G", "H", "I", "M"),
# a factor vector of length 10, with 3 different factor levels
   vec_factor = factor(c("High", "Low", "Medium", "High", "Low", "Medium","High", "Low", "Medium","Low"))
)
# Check the tibble:
problem2_df
```


##### Try to take the mean of each variable in the dataframe

```{r, warning=FALSE}
# Mean of vec_sample:
mean_numeric_sample = mean(pull(problem2_df, vec_sample))
mean_numeric_sample
# Mean of vec_logical:
mean_numeric_logical = mean(pull(problem2_df, vec_logical))
mean_numeric_logical
# Mean of vec_char:
mean_numeric_char = mean(pull(problem2_df, vec_char))
mean_numeric_char 
# Mean of vec_char:
mean_numeric_factor = mean(pull(problem2_df, vec_factor))
mean_numeric_factor
```
*   The result shows that we can successfully calculate the mean of random sample factor(mean `r mean_numeric_sample`) and logical vector(mean `r mean_numeric_logical`). However, for character vector and factor vector, the result will be N/A. 



##### Applies as.numeric function to the logical, characther and factor variables
```{r, results='hide', warning=FALSE}
# Apply for vec_logical and compute mean
   numeric_logical = as.numeric(pull(problem2_df, vec_logical))
# Apply for vec_char and compute mean
   numeric_char = as.numeric(pull(problem2_df, vec_char))
# Apply for vec_factor and compute mean
   numeric_factor = as.numeric(pull(problem2_df, vec_factor))
# Mean of vec_logical:
   mean_numeric_logical_new = mean(numeric_logical)
# Mean of vec_char:
   mean_numeric_char_new = mean(numeric_char)
# Mean of vec_char:
   mean_numeric_factor_new = mean(numeric_factor)
```
*   After converting the variable, we can find that:
   1. The character vector has been changed into a numeric variable but with all equal "NA"
   2. The factor vector has been changed into a numeric variable with 3 different levels as 1 to 3
   3. The logical vector has also been changed into a numeric variable with two levels as 0 or 1
*   The mean can be calculated for logical and factor vectors after converting, but still cannot be calculated for character vector.
*   This shows that factor vectors including characters cannot be calculated directly. Whereas, for logical vectors and character vector, it seems with no differences before and after.
*   I think it is because mean function cannot be applied to vectors containing characters, for example, the even if we have converted character vector, we cannot compute the mean because all in them are "NA". But because logical vector can be translated to 0(FALSE) or 1(TRUE), that can be calculated.

