---
title: "Homework 1"
author: Suning Zhao
Date: September 2022
output: github_document
---

# Problem 1
##### Download the package containing the penguins dataset and load:

```{r}
# Set echo = TRUE as default
knitr::opts_chunk$set(echo = TRUE)
# Open the library of tidyverse
library(tidyverse)
# Load the datafile
data("penguins", package = "palmerpenguins")
# Demonstrate the dataset
head(penguins)
# Compute row, column and mean flipper length
row_penguins = nrow(penguins)
col_penguins = ncol(penguins)
mean_flipper_length = mean(pull(penguins, flipper_length_mm), na.rm = TRUE)
# Compute mean for bill_length_mm, bill_depth_mm
mean_bill_length_mm = mean(pull(penguins, bill_length_mm), na.rm = TRUE)
mean_bill_depth_mm = mean(pull(penguins, bill_depth_mm), na.rm = TRUE)
# Compute mean for body_mass_g, year
mean_body_mass_g = mean(pull(penguins, body_mass_g), na.rm = TRUE)
mean_year = mean(pull(penguins, year), na.rm = TRUE)
```

##### Write a short description of penguins dataset

*   In _penguins_ dataset, there are 8 variables, including species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex and year. Sex is a factor vectors including 2 different values. Species and island are two factor vectors including 3 different values. Bill_length_mm(mean `r mean_bill_length_mm`), bill_depth_mm(mean `r mean_bill_depth_mm`) are two numeric variables. Flipper_length_mm(mean `r  mean_flipper_length`), body_mass_g(mean `r mean_body_mass_g`) and year(mean `r mean_year`) are three int variables.


*   The size of _penguins_ is `r row_penguins` rows and `r col_penguins` lines.

*   The mean _flipper_ _length_ is `r mean_flipper_length`mm.



##### Make a scatterplot

```{r}
ggplot(penguins,aes(y = flipper_length_mm, x = bill_length_mm, color = species)) + geom_point()

ggsave("scatter_plot.pdf", height = 4, width = 6)

```

# Problem 2

##### Create a data frame

```{r}
problem2_df = tibble(
# A random sample of size 10 from a standard Normal distribution
   vec_sample = rnorm (10),
# A logical vector indicating whether elements of the sample are greater than 0
   vec_logical = vec_sample > 0,
# A character vector of length 10
   vec_char = c("0", "1", "2", "3", "4","5", "6", "7", "8", "9"),
# a factor vector of length 10, with 3 different factor levels
   vec_factor = factor(c("High", "Low", "Medium", "High", "Low", "Medium","High", "Low", "Medium","Low"))
)
# Check the tibble:
problem2_df
```


##### Try to take the mean of each variable in the dataframe

```{r}
# Mean of vec_sample:
mean_numeric_sample = mean(pull(problem2_df, vec_sample))
# Mean of vec_logical:
mean_numeric_logical = mean(pull(problem2_df, vec_logical))
# Mean of vec_char:
mean_numeric_char = mean(pull(problem2_df, vec_char))
# Mean of vec_char:
mean_numeric_factor = mean(pull(problem2_df, vec_factor))
```
*   The result shows that we can successfully calculate the mean of random sample factor(mean `r mean_numeric_sample`) and logical vector(mean `r mean_numeric_logical`). However, for character vector and factor vector, the result will be N/A. 



##### Applies as.numeric function to the logical, characther and factor variables
```{r, results='hide'}
# Apply for vec_logical and compute mean
   numeric_logical = as.numeric(problem2_df$vec_logical)
# Apply for vec_char and compute mean
   numeric_char = as.numeric(problem2_df$vec_char)
# Apply for vec_factor and compute mean
   numeric_factor = as.numeric(problem2_df$vec_factor)  
# Mean of vec_logical:
   mean_numeric_logical_new = mean(numeric_logical)
# Mean of vec_char:
   mean_numeric_char_new = mean(numeric_char)
# Mean of vec_char:
   mean_numeric_factor_new = mean(numeric_factor)
```
*   After converting the variable, we can find that:
   1. The character vector has been changed into a numeric variable from 0 to 9
   2. The factor vector has been changed into a numeric variable with 3 different levels as 1 to 3
   3. The logical vector has also been changed into a numeric variable with two levels as 0 or 1
*   The mean can be calculated for all of three variables after converting.This shows that character and factor vectors including characters cannot be calculated directly. Whereas, for logical factors, it seems with no differences before and after.
*   I think it is because mean function cannot be applied to vectors containing characters. But because logical vector can be translated to 0(FALSE) or 1(TRUE), that can be calculated.

