---
title: "Homework 1"
author: Suning Zhao
Date: September 2022
output: github_document
---

# Problem 1
##### Download the package containing the penguins dataset and load:

```{r}
# Set echo = TRUE as default
knitr::opts_chunk$set(echo = TRUE)
# Open the library of tidyverse
library(tidyverse)
# Load the datafile
data("penguins", package = "palmerpenguins")
# Demonstrate the dataset
head(penguins)

```

##### Write a short description of penguins dataset

In _penguins_ dataset, there are 8 variables, including species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex and year.

The size of _penguins_ is `r nrow(penguins)` rows and `r ncol(penguins)` lines.

The mean flipper length is `r mean(pull(penguins, flipper_length_mm), na.rm=TRUE)`mm.


##### Make a scatterplot

```{r}
ggplot(penguins,aes(y = flipper_length_mm, x = bill_length_mm, color = species)) + geom_point()

ggsave("scatter_plot.pdf", height = 4, width = 6)

```

# Problem 2

##### Create a data frame

```{r}
problem2_df = tibble(
# A random sample of size 10 from a standard Normal distribution
   vec_sample = rnorm (10),
# A logical vector indicating whether elements of the sample are greater than 0
   vec_logical = vec_sample > 0,
# A character vector of length 10
   vec_char = c("0", "1", "2", "3", "4","5", "6", "7", "8", "9"),
# a factor vector of length 10, with 3 different factor levels
   vec_factor = factor(c("High", "Low", "Medium", "High", "Low", "Medium","High", "Low", "Medium","Low"))
)
# Check the tibble:
problem2_df
```

##### Try to take the mean of each variable in the dataframe

```{r}
# Mean of vec_sample:
mean(pull(problem2_df, vec_sample))
# Mean of vec_logical:
mean(pull(problem2_df, vec_logical))
# Mean of vec_char:
mean(pull(problem2_df, vec_char))
# Mean of vec_char:
mean(pull(problem2_df, vec_factor))
```

##### Applies as.numeric function to the logical, characther and factor variables
```{r, eval=FALSE}
# Apply for vec_logical and compute mean
   numeric_logical = as.numeric(problem2_df$vec_logical)
# Apply for vec_char and compute mean
   numeric_char = as.numeric(problem2_df$vec_char)
# Apply for vec_factor and compute mean
   numeric_factor = as.numeric(problem2_df$vec_factor)  
# Mean of vec_logical:
   mean(numeric_logical)
# Mean of vec_char:
   mean(numeric_char)
# Mean of vec_char:
   mean(numeric_factor)
```


